# Local Linear Trend Model: dlm Package {#localLinearTrendDLM}
The *local linear trend model* builds on the [local level model](#localLevel).
It adds a time-varying trend, $\nu_{t}$, that follows a random walk.
As before, we observe _y_, which is the underlying level plus noise.

\begin{eqnarray*}
  y_{t} & = & \mu_{t} + \epsilon_{t}, \qquad \epsilon_{t} \sim N(0, \sigma_{\epsilon}^{2}) \\
  \mu_{t} & = & \mu_{t-1} + \nu_{t-1} + \xi_{t}, \qquad \xi_{t} \sim N(0, \sigma_{\xi}^{2}) \\
  \nu_{t} & = & \nu_{t-1} + \zeta_{t}, \qquad \zeta_{t} \sim N(0, \sigma_{\zeta}^{2})
\end{eqnarray*}

This model has five parameters.

---------------------------  --------------------------------------------
$\sigma_{\epsilon}^{2}$      Variance of observation errors, $\epsilon$
$\sigma_{\xi}^{2}$           Variance of transition errors, $\xi$
$\sigma_{\zeta}^{2}$         Variance of slope errors, $\zeta$
$\mu_{0}$                    Initial level of $\mu$
$\nu_{0}$                    Initial level of $\nu$
---------------------------  --------------------------------------------


## Fitting a Local Linear Trend Model (dlm version)

### Problem {-}
You want to build a local linear trend model of your data,
using the special features of the dlm package.

### Solution {-}
The `dlm` documentation refers to this model as the _linear growth model_.

The `dlm` code for estimating a local linear trend model
begins by defining a function capable of creating the appropriate _dlm_
model object from five parameters.

```{r, eval=FALSE}
buildModPoly2 <- function(v) {
  dV <- exp(v[1])
  dW <- exp(v[2:3])
  m0 <- v[4:5]
  dlmModPoly(order=2, dV=dV, dW=dW, m0=m0)
}
```

Notice that the five model parameters are packed into
one 5-element R vector.

The `dlmMLE` uses our `buildModPoly2` function
to find the maximum likelihood estimates (MLE) of the parameters.
It uses numerical optimization, so always check for convergence.

```{r, eval=FALSE}
varGuess <- var(diff(y), na.rm=TRUE)
mu0Guess <- as.numeric(y[1])
lambda0Guess <- 0.0

parm <- c(log(varGuess), log(varGuess), log(varGuess),
          mu0Guess, lambda0Guess)
mle <- dlmMLE(y, parm=parm, buildModPoly2)

if (mle$convergence != 0) stop(mle$message)
```

From the MLE parameters, we can construct the final model object.

```{r, eval=FALSE}
model <- buildModPoly2(mle$par)
```

The `model` object contains the estimated parameters (among other things).

------ -------------------------------------------------------
`V`    Variance of the observations (scalar)
`W`    Variance of the state variables' error terms (matrix)
`m0`   Initial values of the state variables (vector)
------ -------------------------------------------------------

### Example {-}

```{r, eval=TRUE}
library(dlm)

y <- datasets::Nile

buildModPoly2 <- function(v) {
  dV <- exp(v[1])
  dW <- exp(v[2:3])
  m0 <- v[4:5]
  dlmModPoly(order=2, dV=dV, dW=dW, m0=m0)
}

varGuess <- var(diff(y), na.rm=TRUE)
mu0Guess <- as.numeric(y[1])
lambda0Guess <- 0.0

parm <- c(log(varGuess), log(varGuess), log(varGuess),
          mu0Guess, lambda0Guess)
mle <- dlmMLE(y, parm=parm, buildModPoly2)

if (mle$convergence != 0) stop(mle$message)

model <- buildModPoly2(mle$par)
```

### See Also {-}

## Diagnosing a Local Linear Trend Model (dlm version)

### Problem {-}
After fitting a local linear trend model using dlm,
you want to assess the quality of the model.

### Solution {-}
The `tsdiag` function is a generic function for diagnosing time series models,
and the `dlm` package has an implementation.
It produces useful plots for identifying problems in your model.

The diagnostics are based on the posterior distribution defined by the model,
so call `dlmFilter` first to construct the posterior,
then apply `tsdiag` to the result.

```{r, eval=FALSE}
filt <- dlmFilter(y, model)
tsdiag(filt)
```
### Example {-}

### Discussion {-}

### See Also {-}

## Smoothing With a Local Linear Trend Model (dlm version)

### Problem {-}
After building a local linear trend model using dlm,
you want to [smooth](#smoothingVersusFiltering) the data;
that is, remove the noise component from the entire dataset.

### Solution {-}
The `dlm` package provides a function, `dlmSmooth`,
for smoothing your data based on a model.
If $y$ is your data
and `model` is any model created by `dlm`,
such as the recipes in this monograph,
then this call will compute the smoothed data.

```{r, eval=FALSE}
smooth <- dlmSmooth(y, model)
## smooth$s contains the smoothed values
```

### Example {-}

### Discussion {-}

### See Also {-}

For an example of diagnosing a model built with the `dlm` package,
see [Diagnosing a Regression Model, Fixed Coefficients](#diagnoseRegressionFixed).

For an example of smoothing with the `dlm` package,
see [Smoothing With a Regression Model, Fixed Coefficients](#smoothRegressionFixed).

## Filtering With a Local Linear Trend Model (dlm version)

### Problem {-}
After building a local linear trend model using dlm,
you want to [filter](#smoothingVersusFiltering) your time series data;
that is, you want to remove the noise from the data observed so far.

### Solution {-}

### Example {-}

### Discussion {-}

### See Also {-}

## Forecasting with a Local Linear Trend Model

### Problem {-}
You want to forecast a time series
using a local linear trend model built with dlm.

### Solution {-}

### Example {-}

### Discussion {-}

### See Also {-}

