# Appendices {-}

## Random Walk Model with dlm {#randomWalkWithDLM}

The [Random Walk](#randomWalk) chapter shows how to
create a random walk model using the `StructTS` function.
You can also create a random walk model by using the dlm package,
and this appendix shows how.

### Estimation
The code for estimating parameters is very similar to
the code for estimating the local level model.
The difference is that we force $V$, the variance of the observations,
to be zero.

We define a function, `buildRandomWalk`, that builds a `dlm` model object
from two input parameters, `dW` and `m0`.
The parameters are packed into a single, 2-element vector.

```{r, eval=FALSE}
buildRandomWalk <- function(v) {
  dW <- exp(v[1])
  m0 <- v[2]
  dlmModPoly(order=1, dV=0, dW=dW, m0=m0)
}
```

The function calls the `dlmModPoly` function from `dlm` to create the model object.

We need initial guesses for the model parameters.

```{r, eval=FALSE}
varGuess <- var(diff(y), na.rm=TRUE)
mu0Guess <- as.numeric(y[1])
```

Next we call the `dlmMLE` function to estimate the MLE parameters
using numerical optimzation.
Always check for convergence.

```{r, eval=FALSE}
parm <- c(log(varGuess), mu0Guess)
mle <- dlmMLE(y, parm=parm, buildRandomWalk)
if (mle$convergence != 0) stop(mle$message)
```

From the MLE estimates, we can build the final `dlm` model.

```{r, eval=FALSE}
model <- buildRandomWalk(mle$par)
```

We can extract the estimated parameters from `model`, the returned object.

---------- -----------------------------------
`model$W`  Variance of the random walk errors
`model$m0` Initial level
---------- -----------------------------------

### Example {-}

```{r, eval=TRUE}
library(dlm)

y <- datasets::Nile

buildRandomWalk <- function(v) {
  dW <- exp(v[1])
  m0 <- v[2]
  dlmModPoly(order=1, dV=0, dW=dW, m0=m0)
}

varGuess <- var(diff(y), na.rm=TRUE)
mu0Guess <- as.numeric(y[1])

parm <- c(log(varGuess), mu0Guess)
mle <- dlmMLE(y, parm=parm, buildRandomWalk)
if (mle$convergence != 0) stop("Optimizer did not converge")

model <- buildRandomWalk(mle$par)

cat("Transitional variance:", model$W, "\n",
    "Initial level:", model$m0, "\n")
```
