# TEMPLATE CHAPTER {#TEMPLATE}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(introduction).
If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(introduction).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r TEMPLATE-fig, fig.cap='Here is a TEMPLATE figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:TEMPLATE-fig).
Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:TEMPLATE-tab).

```{r TEMPLATE-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a TEMPLATE table!',
  booktabs = TRUE
)
```

You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

## Fitting an XXX Model

## Smoothing With an XXX Model

## Filtering With an XXX Model

## Diagnosing an XXX Model

## Plotting an XXX Model
