--- 
title: "State Space Recipes in R"
author: "Paul Teetor"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: pteetor/StateSpaceRecipes
description: "This monograph is a collection of recipes for creating state-space models in R."
---

```{r globals, include=FALSE}

#
# Document global definitions
#

suppressPackageStartupMessages({
    library(dlm)
})

knitr::opts_chunk$set(prompt=FALSE, echo=TRUE, comment="##")

# Globals for this document
ALT_COLOR = "red"
ALT_STYLE = "dashed"
```

# Introduction

This monograph is a collection of recipes for creating state-space models in R.
I like the power of state-space models,
and R has several excellent packages for building them.
Unfortunately, it's not quite an "out of the box" technology.
Using any package involves numerous little details,
and unless I used the package very recently,
building a model requires pulling out the package documentation,
reading it all over again, and trying to remember how the parts fit together.
One day I got tired of that,
so I put together these recipes.

This is not a tutorial for state-space models.
For a general introduction to state-space modeling,
I recommend the book by Commandeur and Koopman[^CommandeurKoopman2007].

In these notes, I use the `StructTS` function
to create the simpler models,
and I use the `dlm` package
for more complicated models.
There isn't room here to cover other R packages.
If you're interested in a survey of state-space packaqes for R,
I recommend the excellent review by Tusell[^Tusell2011].

## The StructTS function
R includes a function, `StructTS`, which can quickly and easily estimate
the parameters of simple state-space models
such as the *local level* model or the *local linear trend* model.[^Ripley2002]

`StructTS` is one function in a group of functions
which, together, provide many features of state-space modeling.

Function        Purpose
---------       --------
StructTS        Estimate parameters of a simple state-space model
tsdiag          Plot diagnostics for state-space model
KalmanLike      Calculate parameters' log-likelihood (Gaussian model)
KalmanRun       Filter time series data
tsSmooth        Smooth time series data (calls KalmanSmooth)
KalmanForecast  Forecast time series points from model
makeARIMA       Create state-space model equivalent to ARIMA model

## The dlm package
For the advanced recipes,
I use the `dlm` package originally created by Giovanni Petris.[^Petris2010]
The package is very well documented,
and Petris has even written a book regarding state-space models in general
and the `dlm` package in particular.[^PPC2009]
There is also an overview written by Petris and Petrone[^PetrisPetrone2011]
which discusses several R packages with an emphasis on the `dlm` package.

The package contains many useful functions.
This monograph uses these.

Function     Purpose
------------ --------------------------------------------------
dlmModPoly   Construct polynomial model
dlmModReg    Construct regression model
dlmMLE       Estimate maximum likelihood parameters of model
dlmFilter    Filter a time series
dlmSmooth    Smooth a time series
dlmBSample   Draw from the posterior distribution

The package includes a very cool feature,
which is the ability to "add" models together into a compound model.
That feature is not illustrated here,
but I urge any serious user to study the feature.
It would let you, say, easily combine a regression model with an ARMA model
to create a better model your data.

## The examples
Many recipes includes an example.
The examples are intended to be fully stand-alone,
meaning you can cut and paste them directly into R and watch them run.

All examples use some concrete dataset, typically the Nile River data included with R.
The recipes start by assigning the time series data to variable $y$, like this.
```{r, eval=FALSE, echo=TRUE}
y <- datasets::Nile
```
The subsequent code is written in terms of $y$, not a specific dataset.
My goal was to let you copy the recipe, substitute your data for the Nile River data,
and try the recipe for yourself.

## Online materials
R code examples are available on a
public Github repository.

> https://github.com/pteetor/StateSpaceRecipes
