--- 
title: "State Space Recipes in R"
author: "Paul Teetor"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: pteetor/StateSpaceRecipes
description: "This monograph is a collection of recipes for creating state-space models in R."
---

```{r globals, include=FALSE}

#
# Document global definitions
#

suppressPackageStartupMessages({
  library(magrittr)
  library(tidyverse)
  library(dlm)
})

knitr::opts_chunk$set(prompt=FALSE, echo=TRUE, comment="##")

# Globals for this document
ALT_COLOR = "red"
ALT_STYLE = "dashed"
```

# Introduction

This monograph is a collection of recipes for creating state-space models in R.
I like the power of state-space models,
and R has several excellent packages for building them.
Unfortunately, it's not quite an "out of the box" technology.
Using any package involves numerous little details,
and unless I used the package very recently,
building a model requires pulling out the package documentation,
reading it all over again, and trying to remember how the parts fit together.
One day I got tired of that,
so I put together these recipes.

This is not a tutorial for state-space models.
For a general introduction to state-space modeling,
I recommend the book by Commandeur and Koopman[^CommandeurKoopman2007].

In these notes, I use the `StructTS` function
to create the simpler models,
and I use the `dlm` package
for more complicated models.
There isn't room here to cover other R packages.
If you're interested in a survey of state-space packaqes for R,
I recommend the excellent review by Tusell[^Tusell2011].

## The Software
Packages used in this monograph include

-------------------------  -----------------------------------------------
`dlm`                      State-space modeling
`ggplot2' and `ggfortify`  Plotting time series data
`magrittr`                 Provides the convenient "pipe" operator (`%>%`)
`xts` and `zoo`            Tools for time series data
-------------------------  -----------------------------------------------

### The StructTS function
R includes a function, `StructTS`, which can quickly and easily estimate
the parameters of simple state-space models
such as the *local level* model or the *local linear trend* model.[^Ripley2002]

`StructTS` is one function in a group of functions
which, together, provide many features of state-space modeling.

Function        Purpose
---------       --------
StructTS        Estimate parameters of a simple state-space model
tsdiag          Plot diagnostics for state-space model
KalmanLike      Calculate parameters' log-likelihood (Gaussian model)
KalmanRun       Filter time series data
tsSmooth        Smooth time series data (calls KalmanSmooth)
KalmanForecast  Forecast time series points from model
makeARIMA       Create state-space model equivalent to ARIMA model

### The dlm package
For the advanced recipes,
I use the `dlm` package originally created by Giovanni Petris.[^Petris2010]
The package is very well documented,
and Petris has even written a book regarding state-space models in general
and the `dlm` package in particular.[^PPC2009]
There is also an overview written by Petris and Petrone[^PetrisPetrone2011]
which discusses several R packages with an emphasis on the `dlm` package.

The package contains many useful functions.
This monograph uses these.

Function     Purpose
------------ --------------------------------------------------
dlmModPoly   Construct polynomial model
dlmModReg    Construct regression model
dlmMLE       Estimate maximum likelihood parameters of model
dlmFilter    Filter a time series
dlmSmooth    Smooth a time series
dlmBSample   Draw from the posterior distribution

The package includes a very cool feature,
which is the ability to "add" models together into a compound model.
That feature is not illustrated here,
but I urge any serious user to study the feature.
It would let you, say, easily combine a regression model with an ARMA model
to create a better model your data.

## The Examples
Many recipes includes an example.
The examples are intended to be fully stand-alone,
meaning you can cut and paste them directly into R and watch them run.

All examples use some concrete dataset, typically the Nile River data included with R.
Many recipes start by assigning the time series data to variable $y$, like this.

```{r, eval=FALSE, echo=TRUE}
y <- datasets::Nile
```

The subsequent code is written in terms of $y$, not a specific dataset.
My goal was to let you copy the recipe, substitute your data for the Nile River data,
and try the recipe for yourself.

## Smoothing Versus Filtering {#smoothingVersusFiltering}
Every state-space model identifies a noise component in your data,
and there are smoothing and filtering algorithms that
can remove the noise.
The difference between those algorithms is what they assume.

Suppose you have a time series with observations
at times $t_0$ through $t_T$,

$$t_0, t_1, ..., t_T$$

and you want to remove the noise from some intermediate point, say $t_i$.

$$t_0, t_1, ..., t_i, ..., t_T$$

*Smoothing* assumes the entire time series data is available to you,
start to finish, from $t_0$ through $t_T$.

*Filtering* assumes only the data from $t_0$ through $t_i$
is available;
that is, you've not yet seen the data from $t_i+1$ through $t_T$.

Because smoothing has the entire dataset available,
it can estimate the best model and, hence,
does the best job of separating the signal from the noise.
Filtering can only estimate the best "so far",
given what we've observed,
so it estimates an imperfect model and can mistake
some signal for noise (and vice versa).

The practical impact is that smoothing is possible
when analyzing historical data
but impossible for "real-time" data,
that is, data you're observing moment by moment, day by day,
week by week, and so forth.
For the real-time data, you are limited to filtering.
When more data arrives subsequently, the model estimates
may improve and, in fact, the filtered results may change.

## Plotting Time Series Data {#plotting}

### Base graphics
Base R includes a `plot` function for time series data.
The function is quick-and-easy,
but I find it awkward to adjust the plot's appearance.

```{r}
y <- datasets::Nile
plot(y)
```

### ggplot2 with `autoplot`
Use the `autoplot` function of the ggfortify package
to plot your time series data using ggplot2,
the popular graphics package.

```{r}
library(ggfortify)

y <- datasets::Nile
autoplot(y)
```

The result of `autoplot` is a ggplot2 object,
so you can easily customize the plot using ggplot2 functions.

```{r}
autoplot(y, ts.colour = "red") +
  ggtitle("Nile river data") +
  xlab("Year") + ylab("Annual flow") +
  theme_linedraw()
```

### Plotting multivariate time series
Both `plot` and `autoplot` handle multivariate time series data.
We can use the `airquality` built-in dataset to illustrate
(after converting it to time series data).

```{r}
y <- as.ts(airquality[ , 1:4])
```

```{r}
plot(y)
```

```{r}
autoplot(y)
```

## Online materials
R code examples are available in a public Github repository.

> https://github.com/pteetor/StateSpaceRecipes
